---
title: "Data Tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(readr)
```

# Important notes
* Cntrl + Alt + i to get a code block into Rmarkdown
* Using read_csv from readr because it's more efficient and does better job of guessing column types
* shortcut for pipes: Ctrl + Shift + M
* Pipes are good for cutting out dataframes created for intermediary steps
* Use drop down in Environment pane to quickly see variables (instead of str)
* General rule: If column exists, then can write unquoted. However, if you're creating a column you need to name them within quotes (are assigning a string)
* filter is the tidyverse version of subset (structured correctly to work in tidyr syntax)


```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

# Data manipulation
```{r}
catch_data <- catch_original %>% 
  #select(Region, Year, Chinook, Coho, Pink, Chum, Sockeye)
  select(-All, -notesRegCode)
head(catch_data)
```

# Data examination
Data examination for why Chinook is a character variable

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I","1", Chinook)) %>%
  mutate(Chinook = as.numeric(Chinook))
```

Finding spurious characters in Chinook column
```{r}
i <- which(is.na(catch_clean$Chinook))
i
catch_original[i,]
```

# Move from wide to long format
* (pivot_longer now more used that melt, etc commands)
* Can list all that need to be switched (but can be long) so sometimes easier to just say what you don't want converted
* Can write code where use order of column (Chinnok:Pink or 3:5), but then are relying on the fact that the column order doesn't change. Better code written a way that column order doesn't matter.
```{r}
catch_long <- catch_clean %>%
  pivot_longer(cols = -c(Region, Year), names_to = "Species", values_to = "Catch") 
  #below not currently working
  #rename(catch_thousands = Catch) %>% 
  #mutate(catch = Catch_thousands * 1000)

head(catch_long)
```

Put it back in wide format
```{r}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = Region, values_from = Catch)
head(catch_wide)
```

# Summarize data
```{r}
mean_region <- catch_long %>% 
  filter(Species == "Sockeye") %>% 
  group_by(Region, Species) %>% 
  summarise(catch_mean = mean(Catch),
            n=n(), .groups="drop")

#To get rid of the note within R console box we added the code .groups="drop"

head(mean_region)
```

# Join data
Let's do a left join to give full name of region abbreviation
```{r}
region_defs <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
  select(code, mgmtArea)
```

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))
```
The number of rows for a joined table should be the same. Look at your new table and check these things

```{r}
nrow(catch_joined) == nrow(catch_long)
head(catch_joined)
```


