---
title: "Making_a_map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggmap)
library(dplyr)
```

# Read in the shapefile
```{r}
## Read in shapefile using sf
ak_regions <- read_sf("shapefiles/ak_regions_simp.shp")

#this creates a plot for each column in dataframe (region_id, region, and mgmt_area)
plot(ak_regions)  
```

Notice on the plot there's some weird plotting to the right
It's splitting over 180o, so plots the western most Alutians to the right/
We need to define a coordinate reference system (crs)
There is a very nice package crsuggestthat can help you choosing the right projection for your data.

# Change the projection
These data are in WGS_1984 (#4326). 
We're going to transform them to Alaska Albers (EPSG code = 3338).
```{r}
ak_regions_3338 <- ak_regions %>%
  st_transform(crs = 3338)

#Check the reference system
st_crs(ak_regions_3338)
```
Now our plots look better
```{r}
plot(ak_regions_3338)
```

# Grouping our data
```{r}
pop <- read.csv("shapefiles/alaska_population.csv")
head(pop)
```

Notice that the population file doesn't give us the region name.
We need to figure out which region each city is in
Lots of ways to join in the st package (left, where lines cross, etc)

But first we need to coerce the pop dataframe into a sf type
Give it the columns that are coordinates and the datum the data are. (Not what you want them to be! We are assuming 4326; be good with your metadata!).
The remove=F ensures that when we create the geometry column (spatial column) we can retain the original lat and long columns

```{r}
pop_4326 <- st_as_sf(pop, 
                  coords = c('lng', 'lat'),
                  crs = 4326,
                  remove = F)

head(pop_4326)
```

Now our population file is in datum 4326, but the mapping files are in 3338
We need to make them the same
```{r}
pop_3338 <- pop_4326 %>% 
  st_transform(crs=3338)
```

Join the data
```{r}
pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within)

head(pop_joined)
```
# Summarize the data
We can drop the geometry on the population file now (it's sticky and will stick around unless we tell it to drop it)
Here we will make it unsticky (will drop when we summarize the data)
```{r}
pop_region <- pop_joined %>% 
  as.data.frame()
names(pop_region)
```

Now let's join them
```{r}
pop_region <- pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop = sum(population))

head(pop_region)
```

Now we need to join the population information with our region boundary information. Note that now we have a 4th map (total_pop)
```{r}
pop_region_3338 <- left_join(ak_regions_3338, pop_region, by = "region")
plot(pop_region_3338)
```

You can group by and summarize by sf objects as well dataframe
For example, management areas. Let's calcualte population by mgmt area
```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop = sum(total_pop))

plot(pop_mgmt_3338)

```

There are special sf arguements for dplyr. For example, adding the do_union command gets rid of the little gaps due to low resolution maps being grouped
```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop = sum(total_pop), do_union=FALSE)

plot(pop_mgmt_3338)

```

# Making maps
Can use ggplot -it's now integrated with sf objects
Note that order is backwards, so include the arguments mapping.
To make the legend a gradiant use the scale_fill_contiuous command.
```{r}
ggplot() +
  geom_sf(data= pop_mgmt_3338, mapping=aes(fill = total_pop)) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

Now let's add the rivers & city location data:
```{r}
#Bring in rivers data
rivers_3338 <- read_sf("shapefiles/ak_rivers_simp.shp")
st_crs(rivers_3338)

#Plot
#If we don't scale the river data is looks horrible. Use the scale_size command
ggplot() +
  geom_sf(data = pop_region_3338, aes(fill = total_pop)) +
  geom_sf(data = rivers_3338, aes(size = StrOrder), color = "black") +
  geom_sf(data = pop_3338, aes(), size = 0.5) +
  scale_size(range = c(0.01, 0.2), guide = "none") +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high =  "firebrick", labels = comma)


```
# Using base maps to plot our points (like your sites)
Working with another projection, cuz base data uses it and we're stuck with it
```{r}
pop_3857 <- pop_3338 %>%
  st_transform(crs = 3857)
```

Get base map using a function from the ggplot server using a bounding box
The tile servers have some issues with bounding box datums
So you need to run this solution (based on solution found on githup)
Run it on the ggmap object (tiles back from the server)
```{r}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```

Get bounding box
Start by guessing the zoom level (may need to play with this)
Run the fixing function from above
```{r}
bbox <- c(-170, 52, -130, 64)   # This is roughly southern Alaska
ak_map <- get_stamenmap(bbox, zoom = 4)
ak_map_3857 <- ggmap_bbox_to_3857(ak_map)
```

Map the population data
Need background aes stuff to not run, so set inherit.aes = F
Set scale color for legend
```{r}
ggmap(ak_map_3857) + 
  geom_sf(data = pop_3857, aes(color = population), inherit.aes = F) +
  scale_color_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

This way uses ggmap without API keys. There are other ways to map that need the keys

We won't have time to do leaflet coding, but check out the book. It can be really useful.