---
title: "data visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```
TryCatch tries something: if errors print a message, otherwise go get it from the url
```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
    read.csv("data/escapement.csv"),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"))
        return(esc)
    }
)

head(esc)
```

Use dplyr and tidyr skills to calculate annual escapement by species and region
We need to use seperate because don't have year in here
```{r}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("Year", "Month", "Day"), sep="-") %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Species, SASAP.Region, Year) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Sockeye", "Chinook", "Coho", "Pink", "Chum"))
```

Plot these data
```{r}
ggplot(annual_esc, mapping = aes(x=Species, y=escapement, fill=SASAP.Region)) +
  geom_col()
```

Plot these data only for Kodiak
```{r}
kodiak_esc <- annual_esc %>% 
  filter(SASAP.Region=="Kodiak")
```
```{r}
my_theme <- theme_bw() +
  theme(legend.position="bottom")

#The scale command is using the scales package. It works great for log and % scales.
ggplot(kodiak_esc, aes(x=Year, y=escapement, color=Species)) + 
  geom_point()+
  geom_line()+
  ylab("Escapement (num of fish)")+
  ggtitle("Kodiak Escapement")+
  scale_y_continuous(labels=comma)
  my_theme
  
```

Leaflet works great for creating multiple plots. see Section 9.2.4
Creates a clickable map!!!
Distinct only maps unique combinations (if you have many rows with same cordinates)
```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
```

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~Location)
```

